---
title: "Kmer_sample_prepare"
author: "Weier Guo"
date: "6/15/2022"
output: html_document
---

## Introduction ##
After selecting 3 phenotypes - Leaf_indent_width, Leaf_indent_depth, mean vessel diameter (mvd) - for K-mer analysis, we generate 80 samples for each phenotype (40 bulk1 + 40 bulk2).  
After discussion with Isabelle, we decide to collect 38 samples for each bulk (because some samples are missing in the freezer).  
Add freezer Plate, Plate well, and dilution info for every sample.  
Build a bulk table for pooling samples for re-sequencing.  

## Working Processes ##
### import library ###
```{r}
library(tidyverse)
```

### import datasets ###
```{r}
samples <- read_csv("../output/Unique_wdm_samples_20220615.csv")
plate <- read_csv("../input/Seq_plate_info_20220615.csv")
dilute <- read_csv("../input/Plate_dilution_info_20220615.csv")
samples
plate
dilute
```

### Add Plate name into sample file ###
```{r}
samples.plate <- samples %>%
  filter(is.na(Seq.Name1) == FALSE) %>%
  mutate(Plate1 = str_replace(string = Seq.Name1, pattern = "(.*)_(.*)", replacement = "\\1"))
samples.plate$Plate2 = NA
samples.plate$Plate3 = NA
for (i in 1:length(samples.plate$Number)) {
  if (is.na(samples.plate$Seq.Name2[i]) == FALSE) {
    samples.plate$Plate2[i] = str_replace(samples.plate$Seq.Name2[i], "(.*)_(.*)", "\\1")
  }
  if (is.na(samples.plate$Seq.Name3[i]) == FALSE) {
    samples.plate$Plate3[i] = str_replace(samples.plate$Seq.Name3[i], "(.*)_(.*)", "\\1")
  }
}
samples.plate
```

### add plate well ###
```{r}
samples.plate.well <- left_join(samples.plate, plate, by = c("Seq.Name1"="Seq.Name"))
samples.plate.well <- left_join(samples.plate.well, plate, by = c("Seq.Name2"="Seq.Name"))
samples.plate.well <- left_join(samples.plate.well, plate, by = c("Seq.Name3"="Seq.Name"))
samples.plate.well <- samples.plate.well %>%
  mutate(Old.Name = Old.Name.x, Plate.Well1 = Plate.Well.x, Plate.Well2 = Plate.Well.y, Plate.Well3 = Plate.Well) %>%
  select(Number,Old.Name, Group.width, Group.depth, Group.mvd, Seq.Name1, Plate1, Plate.Well1, Seq.Name2, Plate2, Plate.Well2, Seq.Name3, Plate3, Plate.Well3)

samples.plate.well
```

### add dilution info ###
```{r}
#edit dilute data
dilu <- dilute %>%
  mutate(Dilution.lib.H2O = `Dilution (lib+H2O)`) %>%
  select(Libs, Dilution.lib.H2O)
dilu

#add dilution info into dataset and clean up
samples.plate.well.dilu <- left_join(samples.plate.well, dilu, by = c("Plate1"="Libs"))
samples.plate.well.dilu <- left_join(samples.plate.well.dilu, dilu, by = c("Plate2"="Libs"))
samples.plate.well.dilu <- left_join(samples.plate.well.dilu, dilu, by = c("Plate3"="Libs"))
samples.plate.well.dilu <- samples.plate.well.dilu %>%
  mutate(Dilu.lib.H2O1 = Dilution.lib.H2O.x, Dilu.lib.H2O2 = Dilution.lib.H2O.y, Dilu.lib.H2O3 = Dilution.lib.H2O) %>%
  select(Old.Name, Group.width, Group.depth, Group.mvd, Seq.Name1, Plate1, Plate.Well1, Dilu.lib.H2O1, Seq.Name2, Plate2, Plate.Well2, Dilu.lib.H2O2, Seq.Name3, Plate3, Plate.Well3, Dilu.lib.H2O3)
samples.plate.well.dilu
```

## Ouput ##
```{r}
write.csv(samples.plate.well.dilu, "../output/Samples_prepare_pool_20220615.csv")
```

## add barcodes ##
### load resequence files ###
```{r}
seq.samples <- read_csv("../input/Samples_reseq_20220621.csv")
seq.samples <- seq.samples %>%
  select(Old.Name, Group.width, Group.depth, Group.mvd, Seq.Name1, Plate1, Plate.Well1)
seq.samples
```

### import index file ###
```{r}
barcode <- read_csv("../input/pop25_31_barcodes_20220621.csv")
barcode
```

### join two files ###
```{r}
samples.barcode <- left_join(seq.samples, barcode, by=c("Seq.Name1"="Library")) 
samples.barcode <- samples.barcode %>%
  select(Old.Name, Group.width, Group.depth, Group.mvd, Seq.Name1, Plate1, Plate.Well1, `Index 1 / Barcode 1 #`, `Index 2 / Barcode 2 #`, `Barcode / Index 1 seq`, `Index 2 seq`)
samples.barcode
```

### check uniqueness of barcodes ###
```{r}
all.index <- c()
for (i in 1:171) {
  i1 <- as.character(samples.barcode$`Index 1 / Barcode 1 #`[i])
  i2 <- as.character(samples.barcode$`Index 2 / Barcode 2 #`[i])
  i12 <- paste(i1,i2,sep = "_")
  all.index <- c(all.index, i12)
}
length(unique(all.index))
```

### output file with index and barcode ###
```{r}
write_csv(samples.barcode, "../output/Samples_barcoded_20220621.csv")
```


