---
title: "Pheno_selection"
author: "Weier Guo"
date: "5/4/2022"
output: html_document
---

## Introduction 
Investigate 68 phenotypes to select segregated ones for 500 individuals. 

## Working Process ##
### Import Data ###
#### library ####
```{r}
library(tidyverse)
library(ggplot2)
```

#### dataset ####
**biomass headers**
```{r}
header <- readLines("../input/pheno_output_table_blup_20220428.txt",n=1)
header <- header %>%
  str_split("\t") %>%
  unlist()
```

**vessel headers**
```{r}
header.vessel <- readLines("../input/pheno_vessel_20220522.csv",n=1)
header.vessel <- header.vessel %>%
  str_split(",") %>%
  unlist()
```

**biomass data import**
```{r}
pheno <- read_tsv("../input/pheno_output_table_blup_20220428.txt",col_names = header,comment = "Old")
pheno
```

**vessel data import**
```{r}
vessel <- read_csv("../input/pheno_vessel_20220522.csv",col_names = header.vessel,comment = "Geno")
colnames(vessel) <- c("Old.Name","BLUP_tree_height","BLUP_mvd","BLUP_cmvd","BLUP_cvf","BLUP_vgi")
vessel
```

**Cross Type**
```{r}
headerTP <- readLines("../input/XXX_GWR_IFG_FOrWeier.txt",n=1)
headerTP <- headerTP %>%
  str_split("\t") %>%
  unlist()

crosstype <- read_tsv("../input/XXX_GWR_IFG_FOrWeier.txt", col_names = headerTP, comment = "Old") %>%
  select(Old.Name,CrossType)
crosstype
```

### select only GWR genotypes ###
**biomass**
```{r}
pheno <- left_join(pheno,crosstype, by="Old.Name")
pheno
```
**select only GWR genotype in vessel dataset**
```{r}
vessel <- left_join(vessel,crosstype,by = "Old.Name")
vessel
```

### Remove IFG samples ###
**biomass**
```{r}
pheno <- pheno %>%
  filter(CrossType == "GWR") 
pheno
```
**remove IFG genotypes in vessel dataset**
```{r}
vessel <- vessel %>%
  filter(CrossType == "GWR") 
vessel
```

### plot all phenotypes ###
**biomass**
```{r}
pl <- ggplot(pheno,aes(x = pheno$Blup_Transfo_AUC_color_y1_y2_y3)) +
  geom_histogram() +
  ggtitle(colnames(pheno$Blup_Transfo_AUC_color_y1_y2_y3)) 
pl
ggsave("../output/BiomassPheno1_20220511_2.png")
```

**plot pheno distribution in vessel dataset**
```{r}
pl <- ggplot(vessel,aes(x = vessel$BLUP_vgi)) +
  geom_histogram() +
  ggtitle(colnames(vessel$BLUP_vgi)) 
pl
ggsave("../output/VesselPheno5_20220522.png")
```

### Sort phenotypes with less NA to more NA ###
**biomass**
```{r}
pheno.sum <- as.data.frame(matrix(header[-1], nrow = 68))
data.n <- c()
for (i in 2:69) {
  noNA <- !is.na(pheno[[i]])
  noNA <- noNA[noNA == TRUE]
  data.n <- c(data.n, length(noNA))
}
pheno.sum <- cbind(pheno.sum,data.n) %>%
  arrange(desc(data.n))
pheno.sum
```
**vessel**
```{r}
vessel.sum <- as.data.frame(matrix(header.vessel[-1], nrow = 5))
data.n <- c()
for (i in 2:6) {
  noNA <- !is.na(vessel[[i]])
  noNA <- noNA[noNA == TRUE]
  data.n <- c(data.n, length(noNA))
}
vessel.sum <- cbind(vessel.sum,data.n) %>%
  arrange(desc(data.n))
vessel.sum
```
_The results show no NA._

### Check the maximum and minumn sample IDs in candidate phenotypes ###
**biomass**
```{r}
pheno <- pheno %>%
  arrange(desc(Blup_Transfo_time_serie_color_y1)) %>%
  select(Old.Name,Blup_Transfo_time_serie_color_y1, everything())
pheno
```
**vessel**
```{r}
vessel <- vessel %>%
  arrange(desc(BLUP_tree_height)) %>%
  select(Old.Name,BLUP_tree_height, everything())
vessel
```

### Establish bulked samples ###
**biomass**
```{r}
pheno1 <- pheno
bulk1 <- pheno1 %>%
  filter(Old.Name == Old.Name[1:40]) %>%
  select(Old.Name,Blup_Transfo_time_serie_color_y1)
bulk1
```

```{r}
pheno2 <- pheno %>%
  arrange(Blup_Transfo_time_serie_color_y1) 

bulk2 <- pheno2 %>%
  filter(Old.Name == Old.Name[1:40]) %>%
  select(Old.Name, Blup_Transfo_time_serie_color_y1)
bulk2
```

**vessel**
```{r}
vessel1 <- vessel
bulk1_height <- vessel1 %>%
  filter(Old.Name == Old.Name[1:40]) %>%
  select(Old.Name, BLUP_tree_height)
bulk1_height
```

```{r}
vessel2 <- vessel %>%
  arrange(BLUP_tree_height) 

bulk2_height <- vessel2 %>%
  filter(Old.Name == Old.Name[1:40]) %>%
  select(Old.Name, BLUP_tree_height)
bulk2_height
```

### combine bulk1 and bulk2
**biomass**
```{r}
bulk1 <- bulk1 %>%
  mutate(Group = 1)

bulk2 <- bulk2 %>%
  mutate(Group = 2)

Bulk <- full_join(bulk1, bulk2, by=c("Old.Name","Blup_Transfo_time_serie_color_y1","Group"))
Bulk
```
**vessel**
```{r}
bulk1_height <- bulk1_height %>%
  mutate(Group = 1)

bulk2_height <- bulk2_height %>%
  mutate(Group = 2)

Bulk_height <- full_join(bulk1_height, bulk2_height,by=c("Old.Name","BLUP_tree_height","Group"))
Bulk_height
```

### Check bulk1 and bulk2 differences using t-test ###
#### Are the data from each of the 2 groups follow a normal distribution? ####

Null hypothesis: the data are normally distributed  
Alternative hypothesis: the data are not normally distributed  
use the functions with() and shapiro.test() to compute Shapiro-Wilk test for each group of samples.  
```{r}
# Shapiro-Wilk normality test for Bulk1
with(Bulk, shapiro.test(Blup_Transfo_time_serie_color_y1[Group == "1"]))
# Shapiro-Wilk normality test for Bulk2
with(Bulk, shapiro.test(Blup_Transfo_time_serie_color_y1[Group == "2"])) 
```
#### Do the two populations have the same variances? ####
```{r}
res.ftest <- var.test(Blup_Transfo_time_serie_color_y1 ~ Group, data = Bulk)
res.ftest
```

#### Compute unpaired two-samples t-test ####
```{r}
res <- t.test(Blup_Transfo_time_serie_color_y1 ~ Group, data = Bulk)
res
```


### save selected GWR files ###
```{r}
write.csv(pheno,"../input/Pheno_pre_GWR_20220506.csv")
```


