---
title: "F1_genetic_map_Part2"
author: "Weier Guo"
date: '2022-12-05'
output: html_document
---

# Introduction #
Order markers within chromosomes (Continued with Part 1 R Markdown). 

# Outline #
1. Load packages and dataset
2. Switch/Delete alleles (based on marker ordering within chromosome)
3. Redo linkage group 
4. Order markers within each chromosome

# Load dataset and library 
## library 
```{r}
library("tidyverse")
library("qtl")
library("ggplot2")
library("formattable")
library("emmeans")
```

## dataset ##
```{r}
load("../F1_450/New_LinkageGroups/Pdeltoides_GeneticMap_Part1_20221206.RData")
```

# Order Markers within chromosome 
## genetic map plot 
```{r}
plotMap(Pdeltoides_gmap)
```


## Chr12 
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 12)
pull.map(Pdeltoides_gmap, chr = 12)
```

```{r}
rip12 = ripple(Pdeltoides_gmap, chr=12, window = 7)
summary(rip12)
```

```{r}
rip12lik = ripple(Pdeltoides_gmap, chr=12, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip12lik)
```

no change
```{r}
rip12lik0 = switch.order(Pdeltoides_gmap, chr=12, c(1:7), error.prob = 0.005)
pull.map(rip12lik0, chr=12)
```


```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=12, c(1:7), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=12)
```

## Chr19
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 19)
pull.map(Pdeltoides_gmap, chr = 19)
```

```{r}
rip19 = ripple(Pdeltoides_gmap, chr=19, window = 7)
summary(rip19)
```

```{r}
rip19lik = ripple(Pdeltoides_gmap, chr=19, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip19lik)
```

switch (M8,M7) increase LOD by 0, decrease chromosome length by 0.1 original order.
Original Order.
```{r}
rip19lik0 = switch.order(Pdeltoides_gmap, chr=19, c(1:8), error.prob = 0.005)
pull.map(rip19lik0, chr=19)
```

(M8,M7)
```{r}
rip19lik1 = switch.order(Pdeltoides_gmap, chr=19, c(1:6,8,7), error.prob = 0.005)
pull.map(rip19lik1, chr=19)
```

Choose order 1. 
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=19, c(1:6,8,7), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=19)
```

## Chr17
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 17)
pull.map(Pdeltoides_gmap, chr = 17)
```

```{r}
rip17 = ripple(Pdeltoides_gmap, chr=17, window = 7)
summary(rip17)
```


```{r}
rip17lik = ripple(Pdeltoides_gmap, chr=17, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip17lik)
```

original order.
```{r}
rip17lik0 = switch.order(Pdeltoides_gmap, chr=17, c(1:13), error.prob = 0.005)
pull.map(rip17lik0, chr=17)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=17, c(1:13), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=17)
```


## Chr16
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 16)
pull.map(Pdeltoides_gmap, chr = 16)
```

```{r}
rip16 = ripple(Pdeltoides_gmap, chr=16, window = 7)
summary(rip16)
```

```{r}
rip16lik = ripple(Pdeltoides_gmap, chr=16, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip16lik)
```

original order. 
```{r}
rip16lik0 = switch.order(Pdeltoides_gmap, chr=16, c(1:13), error.prob = 0.005)
pull.map(rip16lik0, chr=16)
```

no change. 
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=16, c(1:13), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=16)
```

## Chr18
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 18)
pull.map(Pdeltoides_gmap, chr = 18)
```

```{r}
rip18 = ripple(Pdeltoides_gmap, chr=18, window = 7)
summary(rip18)
```

```{r}
rip18lik = ripple(Pdeltoides_gmap, chr=18, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip18lik)
```

original order.
```{r}
rip18lik0 = switch.order(Pdeltoides_gmap, chr=18, c(1:18), error.prob = 0.005)
pull.map(rip18lik0, chr=18)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=18, c(1:18), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=18)
```


## Chr15
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 15)
pull.map(Pdeltoides_gmap, chr = 15)
```

```{r}
rip15 = ripple(Pdeltoides_gmap, chr=15, window = 7)
summary(rip15)
```

```{r}
rip15lik = ripple(Pdeltoides_gmap, chr=15, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip15lik)
```

no change.
```{r}
rip15lik0 = switch.order(Pdeltoides_gmap, chr=15, c(1:19), error.prob = 0.005)
pull.map(rip15lik0, chr=15)
```


```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=15, c(1:19), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=15)
```











## Chr13
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 13)
pull.map(Pdeltoides_gmap, chr = 13)
```

```{r}
rip13 = ripple(Pdeltoides_gmap, chr=13, window = 7)
summary(rip13)
```

```{r}
rip13lik = ripple(Pdeltoides_gmap, chr=13, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip13lik)
```

original order.
```{r}
rip13lik0 = switch.order(Pdeltoides_gmap, chr=13, c(1:19), error.prob = 0.005)
pull.map(rip13lik0, chr=13)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=13, c(1:19), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=13)
```


## Chr11
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 11)
pull.map(Pdeltoides_gmap, chr = 11)
```

```{r}
rip11 = ripple(Pdeltoides_gmap, chr=11, window = 7)
summary(rip11)
```

```{r}
rip11lik = ripple(Pdeltoides_gmap, chr=11, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip11lik)
```

original order.
```{r}
rip11lik0 = switch.order(Pdeltoides_gmap, chr=11, c(1:20), error.prob = 0.005)
pull.map(rip11lik0, chr=11)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=11, c(1:20), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=11)
```


## Chr14
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 14)
pull.map(Pdeltoides_gmap, chr = 14)
```


```{r}
rip14 = ripple(Pdeltoides_gmap, chr=14, window = 7)
summary(rip14)
```

```{r}
rip14lik = ripple(Pdeltoides_gmap, chr=14, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip14lik)
```

original order. 
```{r}
rip14lik0 = switch.order(Pdeltoides_gmap, chr=14, c(1:22), error.prob = 0.005)
pull.map(rip14lik0, chr=14)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=14, c(1:22), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=14)
```


## Chr7
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 7)
pull.map(Pdeltoides_gmap, chr = 7)
```


```{r}
rip7 = ripple(Pdeltoides_gmap, chr=7, window = 7)
summary(rip7)
```


```{r}
rip7lik = ripple(Pdeltoides_gmap, chr=7, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip7lik)
```

Original order.
```{r}
rip7lik0 = switch.order(Pdeltoides_gmap, chr=7, c(1:22), error.prob = 0.005)
pull.map(rip7lik0, chr=7)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=7, c(1:22), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=7)
```


## Chr9
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 9)
pull.map(Pdeltoides_gmap, chr = 9)
```


```{r}
rip9 = ripple(Pdeltoides_gmap, chr=9, window = 7)
summary(rip9)
```


```{r}
rip9lik = ripple(Pdeltoides_gmap, chr=9, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip9lik)
```

original order. 
```{r}
rip9lik0 = switch.order(Pdeltoides_gmap, chr=9, c(1:25), error.prob = 0.005)
pull.map(rip9lik0, chr=9)
```


no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=9, c(1:25), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=9)
```


## Chr10
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 10)
pull.map(Pdeltoides_gmap, chr = 10)
```

```{r}
rip10 = ripple(Pdeltoides_gmap, chr=10, window = 7)
summary(rip10)
```


```{r}
rip10lik = ripple(Pdeltoides_gmap, chr=10, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip10lik)
```


original order.
```{r}
rip10lik0 = switch.order(Pdeltoides_gmap, chr=10, c(1:29), error.prob = 0.005)
pull.map(rip10lik0, chr=10)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=10, c(1:29), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=10)
```


## Chr4
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 4)
pull.map(Pdeltoides_gmap, chr = 4)
```

```{r}
rip4 = ripple(Pdeltoides_gmap, chr=4, window = 7)
summary(rip4)
```


```{r}
rip4lik = ripple(Pdeltoides_gmap, chr=4, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip4lik)
```

original order.
```{r}
rip4lik0 = switch.order(Pdeltoides_gmap, chr=4, c(1:34), error.prob = 0.005)
pull.map(rip4lik0, chr=4)
```

no change. 
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=4, c(1:34), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=4)
```


## Chr3
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 3)
pull.map(Pdeltoides_gmap, chr = 3)
```


```{r}
rip3 = ripple(Pdeltoides_gmap, chr=3, window = 7)
summary(rip3)
```


```{r}
rip3lik = ripple(Pdeltoides_gmap, chr=3, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip3lik)
```


original order.
```{r}
rip3lik0 = switch.order(Pdeltoides_gmap, chr=3, c(1:34), error.prob = 0.005)
pull.map(rip3lik0, chr=3)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=3, c(1:34), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=3)
```


## Chr5
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 5)
pull.map(Pdeltoides_gmap, chr = 5)
```

```{r}
rip5 = ripple(Pdeltoides_gmap, chr=5, window = 7)
summary(rip5)
```


```{r}
rip5lik = ripple(Pdeltoides_gmap, chr=5, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip5lik)
```

original order.
```{r}
rip5lik0 = switch.order(Pdeltoides_gmap, chr=5, c(1:37), error.prob = 0.005)
pull.map(rip5lik0, chr=5)
```

no change.
```{r}
Pdeltoides_gmap = switch.order(Pdeltoides_gmap, chr=5, c(1:37), error.prob = 0.005)
pull.map(Pdeltoides_gmap, chr=5)
```


## Chr8
```{r}
Pdeltoides_gmap = orderMarkers(Pdeltoides_gmap, chr = 8)
pull.map(Pdeltoides_gmap, chr = 8)
```

```{r}
rip8 = ripple(Pdeltoides_gmap, chr=8, window = 7)
summary(rip8)
```


```{r}
rip8lik = ripple(Pdeltoides_gmap, chr=8, window = 4, method = "likelihood", error.prob = 0.005)
summary(rip8lik)
```

