---
title: "F1_visualization"
author: "Weier Guo"
date: '2022-12-02'
output: html_document
---

# Introduction #
This script is aimed to visualize QTLs for F1 genomic sequenced lines.
The significant QTLs are very narrow after QTL mapping for F1 genomic samples. 
To identify the curve of QTLs, we decide to visualize the observed data. 


## data and library import ##
### library ###
```{r}
library(tidyverse)
library(ggplot2)
```

### observed data ###
```{r}
observed <- read_csv("../F1_450/DND/observed/Obs_PC4_y1_y2.csv") %>%
  as.data.frame()
observed
```


### permutation data ###
```{r}
perm <- read_csv("../F1_450/DND/permutation/Perm_PC4_y1_y2.csv") %>%
  as.data.frame()
perm
```

### import markers ###
```{r}
deltoides.gt <- read_tsv("../F1_450/genotype/Pdeltoides_F1_nigraGeno_20221120.txt") %>%
  as.data.frame()
markers <- colnames(deltoides.gt)[-1]
```


## calculate threshold ##
### multivariate ###
```{r}
# function to collect permutation threshold
perm.threshold <- function(rs, ts) {
  apply(rs, 1, quantile, ts, na.rm = TRUE)
}
```

```{r}
# apply function for each imported permutation results, and collect them
Perm90s <- tibble(Del = c(123),
                  Nig = c(123),
                  Dos = c(123))
Perm95s <- tibble(Del = c(123),
                  Nig = c(123),
                  Dos = c(123))
Perm99s <- tibble(Del = c(123),
                  Nig = c(123),
                  Dos = c(123))
Name <- "PC4_y1_y2"

perm90 <- perm.threshold(perm, 0.90)
Perm90s <- rbind(Perm90s, perm90)
perm95 <- perm.threshold(perm, 0.95)
Perm95s <- rbind(Perm95s, perm95)
perm99 <- perm.threshold(perm, 0.99)
Perm99s <- rbind(Perm99s, perm99)

Perm90s <- Perm90s %>%
  filter(Del != 123)
Perm95s <- Perm95s %>%
  filter(Del != 123)
Perm99s <- Perm99s %>%
  filter(Del != 123)
```

### add phenotype name ###
```{r}
Perm90s.pheno <- cbind(Name, Perm90s)
Perm95s.pheno <- cbind(Name, Perm95s)
Perm99s.pheno <- cbind(Name, Perm99s)
```


## sort markers with QTL t-values ##
```{r}
Del.QTL <- tibble(Markers = markers, Del = observed$Del)
Del.QTL %>%
  arrange(desc(Del))
```


## visualization ##
### t-value visualization ###
```{r}
trans.perm <- t(perm) %>%
  as.data.frame()
ggplot(trans.perm, aes(x=trans.perm[,1])) +
  geom_histogram(bins = 50)
```


### QTL curve ###
```{r}
tibble(marker = 1:nrow(observed), Del = observed$Del) %>%
  ggplot(aes(x = marker, y = Del)) +
  geom_point() +
  geom_hline(yintercept = Perm90s$Del, color="red")
```













