---
title: "QTL_selection"
author: "Weier Guo"
date: '2022-08-26'
output: html_document
---

### Introduction ###
This Notebook is to select QTLs from the Pvalue datasets from [combined QTL]("https://github.com/guoweier/CADVIP/blob/main/QTL/scripts/combined_QTL_20220808.Rmd"). 

Selection criteria:
- 0.05 < p-value < 0.1: suggestive linkage
- 0.01 < p-value < 0.05: significant linkage
- p-value < 0.01: confirmed linkage

### import package and datasets ###
#### package ####
```{r}
library(tidyverse)
```

#### datasets prepared ####
```{r}
p.files <- tibble(filename = list.files("../Pvalue_Leaf_DND/"))
p.files.paths <- p.files %>%
  mutate(filepath = paste("../Pvalue_Leaf_DND/", filename, sep = "")) %>%
  mutate(phenotype = str_replace(string = filename, pattern = "(.*).csv", replacement = "\\1"))
head(p.files.paths)
```

### Recurrently load files and do selection ###
```{r}
Linkage.define <- function(pvalue) {
  if (pvalue <= 0.1 & pvalue >= 0.05) {
    "suggestive"
  }
  else if (pvalue < 0.05 & pvalue >= 0.01) {
    "significant"
  }
  else if (pvalue < 0.01) {
    "confirmed"
  }
}
```



```{r}
Del.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Del_P_adj = c(123))
Nig.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Nig_P_adj = c(123))
Dos.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Dos_P_adj = c(123))

for (i in 1:length(p.files.paths$filename)) {
  p.file <- read_csv(p.files.paths$filepath[i])
  # Deltoides
  Dels <- p.file %>%
    filter(Del_P_adj != "NA" & Del_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Del_P_adj)
  Del.QTLs <- rbind(Del.QTLs,Dels)
  # Nigra
  Nigs <- p.file %>%
    filter(Nig_P_adj != "NA" & Nig_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Nig_P_adj)
  Nig.QTLs <- rbind(Nig.QTLs,Nigs)
  # Dosage
  Doss <- p.file %>%
    filter(Dos_P_adj != "NA" & Dos_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Dos_P_adj)
  Dos.QTLs <- rbind(Dos.QTLs,Doss)
}
```


### adjust QTLs ###
#### P. deltoides ####
```{r}
Del.QTLs.adj <- Del.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Del_P_adj))
Del.QTLs.adj
```

#### P. nigra ####
```{r}
Nig.QTLs.adj <- Nig.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Nig_P_adj))
Nig.QTLs.adj
```

#### Dosage ####
```{r}
Dos.QTLs.adj <- Dos.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Dos_P_adj))
Dos.QTLs.adj
```

#### Dosage and Nigra ####
```{r}
Nig.Dos.QTLs.adj <- Nig.Dos.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(NigDos_P_adj))
Nig.Dos.QTLs.adj
```

### export files ###
```{r}
write_csv(Del.QTLs.adj, "../loci/Pdeltoides_QTL_DND_20220925.csv")
write_csv(Nig.QTLs.adj, "../loci/Pnigra_QTL_DND_20220925.csv")
write_csv(Dos.QTLs.adj, "../loci/Dosage_QTL_DND_20220925.csv")
write_csv(Nig.Dos.QTLs.adj, "../loci/PnigraDosage_QTL_lm_20220918.csv")
```




### Dosage embeded into Nigra genotype ###
#### datasets prepared ####
```{r}
p.files <- tibble(filename = list.files("../Pvalue_Leaf_DelNig/"))
p.files.paths <- p.files %>%
  mutate(filepath = paste("../Pvalue_Leaf_DelNig/", filename, sep = "")) %>%
  mutate(phenotype = str_replace(string = filename, pattern = "(.*).csv", replacement = "\\1"))
head(p.files.paths)
```

#### adjust dataset ####
```{r}
Del.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Del_P_adj = c(123))
Nig.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Nig_P_adj = c(123))
for (i in 1:length(p.files.paths$filename)) {
  p.file <- read_csv(p.files.paths$filepath[i])
  # P.deltoides
  Dels <- p.file %>%
    filter(Del_P_adj != "NA" & Del_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Del_P_adj)
  Del.QTLs <- rbind(Del.QTLs,Dels)
  # P. nigra
  Nigs <- p.file %>%
    filter(Nig_P_adj != "NA" & Nig_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Nig_P_adj)
  Nig.QTLs <- rbind(Nig.QTLs,Nigs)
}
```

#### P. deltoides ####
```{r}
Del.QTLs.adj <- Del.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Del_P_adj))
Del.QTLs.adj
```

#### P. nigra ####
```{r}
Nig.QTLs.adj <- Nig.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Nig_P_adj))
Nig.QTLs.adj
```

### export files ###
```{r}
write_csv(Del.QTLs.adj, "../Pdeltoides_QTL_DelNig_20220829.csv")
write_csv(Nig.QTLs.adj, "../Pnigra_QTL_DelNig_20220829.csv")
```




### Dosage embeded into Nigra genotype ###
#### datasets prepared ####
```{r}
p.files <- tibble(filename = list.files("../Pvalue_Leaf_lmer/"))
p.files.paths <- p.files %>%
  mutate(filepath = paste("../Pvalue_Leaf_lmer/", filename, sep = "")) %>%
  mutate(phenotype = str_replace(string = filename, pattern = "(.*).csv", replacement = "\\1"))
head(p.files.paths)
```

#### adjust dataset ####
```{r}
Del.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Del_P_adj = c(123))
Nig.QTLs <- tibble(Pheno = c("Test"),
                   Marker = c("Test"),
                   Nig_P_adj = c(123))
for (i in 1:length(p.files.paths$filename)) {
  p.file <- read_csv(p.files.paths$filepath[i])
  # P.deltoides
  Dels <- p.file %>%
    filter(Del_P_adj != "NA" & Del_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Del_P_adj)
  Del.QTLs <- rbind(Del.QTLs,Dels)
  # P. nigra
  Nigs <- p.file %>%
    filter(Nig_P_adj != "NA" & Nig_P_adj <= 0.1) %>%
    mutate(Pheno = p.files.paths$phenotype[i]) %>%
    select(Pheno, Marker, Nig_P_adj)
  Nig.QTLs <- rbind(Nig.QTLs,Nigs)
}
```

#### P. deltoides ####
```{r}
Del.QTLs.adj <- Del.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Del_P_adj))
Del.QTLs.adj
```

#### P. nigra ####
```{r}
Nig.QTLs.adj <- Nig.QTLs %>%
  filter(Pheno != "Test") %>%
  rowwise() %>%
  mutate(Linkage_Class = Linkage.define(Nig_P_adj))
Nig.QTLs.adj
```

### export files ###
```{r}
write_csv(Del.QTLs.adj, "../Pdeltoides_QTL_lmer_20220829.csv")
write_csv(Nig.QTLs.adj, "../Pnigra_QTL_lmer_20220829.csv")
```


